EDA

### 2020-02-29 src/draft.Rmd

pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
write.csv(pct_final, here("data","data_final.csv"))

@@

# goodness of fit (all data)
library(broom)
linear1_fit<-augment(linear1)
sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 5.614843
linear2_fit<-augment(linear2)
sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 5.614843

```
linear 1: Multiple R-squared:  0.4796,	Adjusted R-squared:  0.155 
linear 2: Multiple R-squared:  0.4796,	Adjusted R-squared:  0.155

we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400). We ignored the two municipalities and refit the linear model and observed that r^2 improved to 0.4882 (Multiple R-squared:  0.4882,	Adjusted R-squared:  0.1605).
Still not too much difference bewteen the model with and without variable year.

@@

# goodness of fit
fit_linear_test<-augment(linear_test)
sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 5.775885

@@

# goodness of fit
fit_linear_test_2<-augment(linear_test_2)
sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) # 5.775885

```
Question: How to judge goodness of fit? Multiple R square? Adjusted R square? or mean prediction error?
**TODO: create training/test sets to test the prediction power!

@@

>>>>

pct_final<-assessment2017 %>% full_join(assessment2018) %>% full_join(assessment2019)
pct_final$tax<-as.numeric(pct_final$tax)
write.csv(pct_final, here("data","data_final.csv"))

@@

# goodness of fit (all data)
library(broom)
linear1_fit<-augment(linear1)
sqrt(sum((linear1_fit$.resid)^2)/nrow(pct_final)) # 6.672786
linear2_fit<-augment(linear2)
sqrt(sum((linear2_fit$.resid)^2)/nrow(pct_final)) # 7.248359

```
linear 1: Residual standard error: 7.296 on 138 degrees of freedom
Multiple R-squared:  0.265,	Adjusted R-squared:  0.1265 
PMSE = 6.672786

linear 2: Residual standard error: 7.869 on 140 degrees of freedom
Multiple R-squared:  0.1327,	Adjusted R-squared:  -0.01593 
F-statistic: 0.8929 on 24 and 140 DF,  p-value: 0.6112
PMSE = 7.248359

we re-looked at the data and found percentage changes of tax in "Langley - City" and "Langley - Township" are too large (~400). We ignored the two municipalities and refit the linear model

@@

# goodness of fit
fit_linear_test<-augment(linear_test)
sqrt(sum((fit_linear_test$.resid)^2)/nrow(test)) # 6.767493

@@

# goodness of fit
fit_linear_test_2<-augment(linear_test_2)
sqrt(sum((fit_linear_test_2$.resid)^2)/nrow(test)) #  7.50047

```
After filtering Langely
with variable year:
Residual standard error: 7.429 on 122 degrees of freedom
Multiple R-squared:  0.2974,	Adjusted R-squared:  0.1592 
F-statistic: 2.152 on 24 and 122 DF,  p-value: 0.003567
PMSE: 6.767493

without variable year
Residual standard error: 8.167 on 124 degrees of freedom
Multiple R-squared:  0.137,	Adjusted R-squared:  -0.01609 
F-statistic: 0.8949 on 22 and 124 DF,  p-value: 0.6014
PMSE: 7.50047

```{r}
linear3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=pct_final)
summary(linear3)
fit_linear3<-augment(linear3)
sqrt(sum((fit_linear3$.resid)^2)/nrow(test)) #  7.50047

Question: How to judge goodness of fit? Multiple R square? Adjusted R square? or mean prediction error?
**TODO: create training/test sets to test the prediction power!

@@

Prediction power
train/test
```{r}
set.seed(450)
train_ind<-sample(165,120)
train1<-pct_final[train_ind,]
test1<-pct_final[-train_ind,]
```

linear model
```{r}
# full model with year
newx<-test1[,-c(5,6,7,8)]
y<-test1[,c(8)]
linear_1<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
resid<-predict(linear_1,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test1))
# pmse = 6.5688

# full model without year
linear_2<-lm(rate_pct~factor(TaxClassCode)+factor(Municipalities)+assessTotal_pct+tax,data=train1)
resid<-predict(linear_2,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test1))
# 6.855429

# reduced variable
linear_3<-lm(rate_pct~factor(Year)+factor(TaxClassCode)+factor(Municipalities),data=train1)
resid<-predict(linear_3,newdata = newx) - y
sqrt(sum(resid^2)/nrow(test1))
# 6.570848
```
Lasso
```{r}
# create the whole matrix
dummy_year<-dummy(pct_final$Year)
dummy_municipal<-dummy(pct_final$Municipalities)
dummy_taxclass<-dummy(pct_final$TaxClassCode)
# build x matrix
x<-cbind(dummy_year,dummy_municipal)
x<-cbind(x,dummy_taxclass)
x<-cbind(x,as.numeric(pct_final$assessTotal_pct))
x<-cbind(x,as.numeric(pct_final$tax))
y<-as.matrix(pct_final$rate_pct)
dim(x) # 165  29
dim(y)
# creat x_train matrix and y_train
x_train<-x[train_ind,]
y_train<-y[train_ind,]
# create x_test matrix
x_test<-x[-train_ind,]
y_test<-y[-train_ind,]

lambdas <- 10^seq(2, -3, by = -.1)
# Setting alpha = 1 implements lasso regression
set.seed(450)
lasso_reg <- cv.glmnet(x_train, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test)
# 7.012128	
```

Ridge
```{r}
ridge_reg = glmnet(x_train, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
set.seed(450)
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0, lambda = lambdas, nfolds=10)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test)
# 6.885151	
```
